---
title: "P8105_hw1_xs2569"
author: "Xun Sun"
date: 2024-09-12
output: 
  html_document:
    toc: true
    toc_float: true
---

# 1] problem 1
a short description of the penguins dataset

### 1.1 the data in this dataset
including names / values of important variables
```{r}
library(ggplot2)
data("penguins", package = "palmerpenguins")
penguins
summary(penguins)
View(penguins)
names(penguins)  
head(penguins)   

```
There are 8 variables, named species, island, bill_length_mm, bill_depth_mm, flopper_length_mm, body_mass, sex, year.

### 1.2 the size of the dataset
```{r}
nrow_dataset <- nrow(penguins) # Get the number of columns in the dataset
ncol_dataset <- ncol(penguins) # Print the size of the dataset

cat("The dataset has", nrow_dataset, "rows and", ncol_dataset, "columns.\n")
```
### 1.3 the mean flipper length
```{r}
mean_flipper_length <- mean(penguins$flipper_length_mm, na.rm = TRUE)
cat("The mean flipper length is", mean_flipper_length, "\n")
```

### 1.4 a scatterplot 

```{r warning = FALSE}

plot1 <- ggplot(penguins, aes(bill_length_mm, flipper_length_mm,color = species)) + geom_point() + scale_color_manual(values = c("Adelie" = "red", "Chinstrap" = "blue", "Gentoo" = "green"))

plot1

ggsave("penguin_scatterplot.png", plot = plot1, width = 8, height = 6, units = "in")

```
The scatterplot reveals that although there is some overlap in the data among the three penguin species, each has distinct characteristics regarding bill length and flipper length. 
Specifically, the Gentoo species generally exhibits higher flipper lengths compared to the other two species. 
Meanwhile, the Chinstrap and Adelie species show similar flipper lengths, but the Chinstrap species tends to have a larger bill length than the Adelie

# 2] Problem 2 
### 2.1 Create a data frame + take the mean 
```{r}
library(tidyverse)

set.seed(1234)
sd <- tibble(
  normal_dist = rnorm(10),
  is_greaterthan0 = normal_dist > 0,
  char_vec = ("DANCINGCAT"),
  char_vec2 = ("1234567890"),
  factor_vec = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE
)),
  factor_vec2 = factor(sample(c("1", "2", "3"), 10, replace = TRUE
)))

mean_normal_dist <- mean(pull(sd, normal_dist))
mean_is_greaterthan0 <- mean(pull(sd, is_greaterthan0))
### mean_char_vec <- mean(pull(sd, char_vec)) 
# This will give an error
### mean_char_vec2 <- mean(pull(sd, char_vec2)) 
# This will give an error
### mean_factor_vec <- mean(pull(sd, mean_factor_vec)) 
# This will give an error
### mean_factor_vec2 <- mean(pull(sd, mean_factor_vec2)) 
# This will give an error

print(mean_normal_dist)
print(mean_is_greaterthan0)


```
###### random_sample: This works because random_sample is numeric
###### logical_vector: This works because TRUE/FALSE can be coerced to 1/0
###### character_vector: This won't work because it's a character variable (both numbers and strings)
###### factor_vector: This won't work because it's a factor



### 2.2 convert variables + explain why
```{r eval=FALSE}
as_numeric_logical <- as.numeric(pull(sd, is_greaterthan0))
as_numeric_char <- as.numeric(pull(sd, char_vec)) 
as_numeric_char2 <- as.numeric(pull(sd, char_vec2))
as_numeric_factor <- as.numeric(as.character(pull(sd, factor_vec)))
as_numeric_factor2 <- as.numeric(as.character(pull(sd, factor_vec2)))

mean_as_numeric_logical <- mean(as_numeric_logical)
mean_as_numeric_char <- mean(as_numeric_char, na.rm = TRUE) # NA values are ignored
mean_as_numeric_char2 <- mean(as_numeric_char2, na.rm = TRUE) # NA values are ignored
mean_as_numeric_factor <- mean(as_numeric_factor, na.rm = TRUE) # NA values are ignored
mean_as_numeric_factor2 <- mean(as_numeric_factor2, na.rm = TRUE)

print(mean_as_numeric_logical)
print(mean_as_numeric_char)
print(mean_as_numeric_char2)
print(mean_as_numeric_factor)
print(mean_as_numeric_factor2)

```

#### What Happens:
###### logical to Numeric: The logical values (`TRUE`, `FALSE`) are coerced to 1 and 0.
###### character to Numeric: Characters containing letters or strings cannot be directly converted to numeric values, so `as.numeric()` on character variables will result in `NA` values, but characters containing numbers can be converted to numeric numbers.
###### factor to Numeric: Factors have underlying integer representations (their levels). When you coerce a factor to numeric, R converts each level to its corresponding integer code.

##### Does This Help Explain the `mean()` Issue?
###### logical variables: taking the mean works because R converts `TRUE`/`FALSE` to 1 and 0
###### character variables: letters cannot be coerced to numeric, but character numbers can.
###### factor variables, when the level names contain letters, it cannot be coerced to numeric, but factors named by character numbers can.