---
title: "P8105_hw1_xs2569"
author: "Xun Sun"
date: 2024-09-12
output: 
  html_document:
    toc: true
    toc_float: true
---

# 1] problem 1
a short description of the penguins dataset

### 1.1 the data in this dataset
including names / values of important variables
```{r}
library(ggplot2)
data("penguins", package = "palmerpenguins")
penguins
summary(penguins)
View(penguins)
names(penguins)  
head(penguins)   

```
### 1.2 the size of the dataset
```{r}
nrow_dataset <- nrow(penguins) # Get the number of columns in the dataset
ncol_dataset <- ncol(penguins) # Print the size of the dataset

cat("The dataset has", nrow_dataset, "rows and", ncol_dataset, "columns.\n")
```
### 1.3 the mean flipper length
```{r}
mean_flipper_length <- mean(penguins$flipper_length_mm, na.rm = TRUE)
cat("The mean flipper length is", mean_flipper_length, "\n")
```

### 1.4 a scatterplot 

```{r}

plot1 <- ggplot(penguins, aes(bill_length_mm, flipper_length_mm,color = species)) + geom_point() + scale_color_manual(values = c("Adelie" = "red", "Chinstrap" = "blue", "Gentoo" = "green"))

plot1

ggsave("penguin_scatterplot.png", plot = plot1, width = 8, height = 6, units = "in")

```

# 2] Problem 2
```{r}
library(tidyverse)

set.seed(1234)
sd <- tibble(
  normal_dist = rnorm(10),
  is_greaterthan0 = normal_dist > 0,
  char_vec = ("DANCINGCAT"),
  factor_vec = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE
)))

mean_normal_dist <- mean(pull(sd, normal_dist))
mean_is_greaterthan0 <- mean(pull(sd, is_greaterthan0))
### mean_char_vec <- mean(pull(sd, char_vec)) 
# This will give an error
### mean_factor_vec <- mean(pull(sd, mean_factor_vec)) 
# This will give an error

print(mean_normal_dist)
print(mean_is_greaterthan0)

```


```{r eval = FALSE}
as_numeric_logical <- as.numeric(pull(sd, is_greaterthan0))
as_numeric_char <- as.numeric(pull(sd, char_vec)) 
as_numeric_factor <- as.numeric(as.character(pull(sd, factor_vec)))

mean_as_numeric_logical <- mean(as_numeric_logical)
mean_as_numeric_char <- mean(as_numeric_char, na.rm = TRUE) # NA values are ignored
mean_as_numeric_factor <- mean(as_numeric_factor, na.rm = TRUE) # NA values are ignored

print(mean_as_numeric_logical)
print(mean_as_numeric_char)
print(mean_as_numeric_factor)