---
title: "P8105_hw1_xs2569"
author: "Xun Sun"
date: 2024-09-12
output: 
  github_document:
---

# 1] problem 1
a short description of the penguins dataset

### 1.1 the data in this dataset
including names / values of important variables
```{r}
library(ggplot2)
data("penguins", package = "palmerpenguins")
summary(penguins)

```
There are 8 variables, named species, island, bill_length_mm, bill_depth_mm, flopper_length_mm, body_mass, sex, year.

Species [qualitative-nominal] : It includes Adelie, Chinstrap, and Gentoo penguins.

Island [qualitative-nominal] : Observations are from Biscoe, Dream, and Torgersen islands.

Bill Length (mm) [quantitative-continuous] : Ranging from 32.10 to 59.60 mm, with an average of approximately 44 mm.

Bill Depth (mm) [quantitative-continuous] : Ranging from 13.10 to 21.50 mm, with an average of about 17 mm.

Flipper Length (mm) [quantitative-continuous] : Ranging from 172.0 to 231.0 mm, with an average of about 201 mm.

Body Mass (g) [quantitative-continuous] : Ranging from 2700 to 6300 g, with an average of about 4202 g.

Sex [qualitative-binary] : Most records indicate female or male, with 11 unknown.

Year [quantitative-discrete] : Data spans from 2007 to 2009, with a focus on the year 2008.

### 1.2 the size of the dataset
```{r}
nrow_dataset <- nrow(penguins) # Get the number of columns in the dataset
ncol_dataset <- ncol(penguins) # Print the size of the dataset

cat("The dataset has", nrow_dataset, "rows and", ncol_dataset, "columns.\n")
```
### 1.3 the mean flipper length
```{r}
mean_flipper_length <- mean(penguins$flipper_length_mm, na.rm = TRUE)
cat("The mean flipper length is", mean_flipper_length, "\n")
```

### 1.4 a scatterplot 

```{r warning = FALSE}

plot1 <- ggplot(penguins, aes(bill_length_mm, flipper_length_mm,color = species)) + geom_point() + scale_color_manual(values = c("Adelie" = "red", "Chinstrap" = "blue", "Gentoo" = "green"))

plot1

ggsave("penguin_scatterplot.png", plot = plot1, width = 8, height = 6, units = "in")

```
The scatterplot reveals that although there is some overlap in the data among the three penguin species, each has distinct characteristics regarding bill length and flipper length. 
Specifically, the Gentoo species generally exhibits higher flipper lengths compared to the other two species. 
Meanwhile, the Chinstrap and Adelie species show similar flipper lengths, but the Chinstrap species tends to have a larger bill length than the Adelie

# 2] Problem 2 
### 2.1 Create a data frame + take the mean 
```{r}
library(tidyverse)

set.seed(1234)
sd <- tibble(
  normal_dist = rnorm(10),
  is_greaterthan0 = normal_dist > 0,
  char_vec = ("DANCINGCAT"),
  char_vec2 = ("1234567890"),
  factor_vec = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE
)),
  factor_vec2 = factor(sample(c("1", "2", "3"), 10, replace = TRUE
)))

mean_normal_dist <- mean(pull(sd, normal_dist))
mean_is_greaterthan0 <- mean(pull(sd, is_greaterthan0))
### mean_char_vec <- mean(pull(sd, char_vec)) 
# This will give an error
### mean_char_vec2 <- mean(pull(sd, char_vec2)) 
# This will give an error
### mean_factor_vec <- mean(pull(sd, mean_factor_vec)) 
# This will give an error
### mean_factor_vec2 <- mean(pull(sd, mean_factor_vec2)) 
# This will give an error

print(mean_normal_dist)
print(mean_is_greaterthan0)


```
###### random_sample: This works because random_sample is numeric
###### logical_vector: This works because TRUE/FALSE can be coerced to 1/0
###### character_vector: This won't work because it's a character variable (both numbers and strings)
###### factor_vector: This won't work because it's a factor



### 2.2 convert variables + explain why
```{r eval=FALSE}
as_numeric_logical <- as.numeric(pull(sd, is_greaterthan0))
as_numeric_char <- as.numeric(pull(sd, char_vec)) 
as_numeric_char2 <- as.numeric(pull(sd, char_vec2))
as_numeric_factor <- as.numeric(as.character(pull(sd, factor_vec)))
as_numeric_factor2 <- as.numeric(as.character(pull(sd, factor_vec2)))

mean_as_numeric_logical <- mean(as_numeric_logical)
mean_as_numeric_char <- mean(as_numeric_char, na.rm = TRUE) # NA values are ignored
mean_as_numeric_char2 <- mean(as_numeric_char2, na.rm = TRUE) # NA values are ignored
mean_as_numeric_factor <- mean(as_numeric_factor, na.rm = TRUE) # NA values are ignored
mean_as_numeric_factor2 <- mean(as_numeric_factor2, na.rm = TRUE)

print(mean_as_numeric_logical)
print(mean_as_numeric_char)
print(mean_as_numeric_char2)
print(mean_as_numeric_factor)
print(mean_as_numeric_factor2)

```

#### What Happens:
###### logical to Numeric: The logical values (`TRUE`, `FALSE`) are coerced to 1 and 0.
###### character to Numeric: Characters containing letters or strings cannot be directly converted to numeric values, so `as.numeric()` on character variables will result in `NA` values, but characters containing numbers can be converted to numeric numbers.
###### factor to Numeric: when the level names contain letters, it cannot be coerced to numeric, but factors named by character numbers can.

##### Does This Help Explain the `mean()` Issue?
###### logical variables: taking the mean works because R converts `TRUE`/`FALSE` to 1 and 0
###### character variables: letters cannot be coerced to numeric, but character numbers can.
###### factor variables, when the level names contain letters, it cannot be coerced to numeric, but factors named by character numbers can.